---
title: 'Update to "assessing far UVC interventions with an individual based infectious disease model"'
author:
  - Charlie Whittaker^[Imperial College London], Tom Brewer^[Imperial College London], Adam Howes^[Independent]
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{blueprint}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: REFERENCES.bib
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 330,
  out.width = "90%"
)

```

```{r}
library(helios)
library(individual)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
theme_set(theme_minimal())
blueprint_colours <- colorRampPalette(c("#00AFFF", "#03113E"))(4)
exemplar_data <- readRDS("./exemplar_simulations.rds")
parameter_sweep_data <- readRDS("./parameter_sweep_data.rds")
```

# Overview

This document builds on [previous work](https://mrc-ide.github.io/helios/articles/blueprint.html) that introduced `helios`, an individual-based modelling framework for exploring the impact that installation of far UVC could have on the transmission and burden of respiratory infections.
In this document, we present a series of analyses evaluating the impact that installation of far UVC in non-household settings could have on the burden....

# Major Model Changes

Include relevant details about the model here, and changes from the first report:

## The Consistent Use of Data from the United States of America {#locations}

Add some subsections here about the data we now use.
Can basically copy the structure of the first report.

## Location-Specific Variability in "Riskiness"

Previously, we modelled differences in "riskiness" (i.e. how receptive to onwards transmission it is) between different Setting Types (schools, workplaces, leisure settings and households).
However, we made the simplifying assumption that all Locations with a particular Setting Type (e.g. all schools) had the same "riskiness", which is unrealistic.
We have now updated thte model to explicitly incorporate variation in "riskiness" between Locations belonging to the same Setting Type.
What follows below is a mathematical description of how this addition to the model works, how it was derived, and how we have parameterised this for the purposes of the new results presented in this report.

### Updating the Force of Infection to Include Location-Specific Variation in Riskiness/Receptivity to Transmission

As a reminder, the previous version of the force of infection experienced by individual $i$ in Location $m$ at time $t$ was calculated as follows:

$$ \lambda_{i,j,m}(t) = \frac{\beta_{m} I_{j,m}(t)}{N_{j,m}} $$

where $I_m(t)$ are the number of infectious individuals in Location $j$ of Setting Type $m$ at time $t$ , $N_{j,m}$ are the total number of individuals visiting Location $j$ of Setting Type $m$ at time $t$ and $\beta_m$ is a Setting Type-specific (i.e. workplace, school, household, leisure-venue) transmissibility parameter that models **on average** how much transmission happens in a particular Setting Type (e.g. workplaces) compared to other Settings Types (e.g. schools).
The total force of infection $\lambda_T$ experienced across all Setting Types is then summed and forms the basis for calculating an individual's probability of infection during one time-step $t$:

$$ P(infection) = 1 - e^{- \lambda_{i,T}t}. $$

Previously, $\beta$ was specific to the Setting Type (i.e. it was different for schools, workplaces, households and leisure settings) but assumed to be the same same for every location belonging to a particular setting type - that is, all individual locations of a particular Setting Type (e.g. all schools) had the same $\beta$.
We have now relaxed this assumption.
The updated force of infection term is now the following:

$$ \lambda_{i,j,m}(t) = \frac{\beta_{j,m} I_{j,m}(t)}{N_{j,m}} $$

where:\
$$ \beta_{j,m} = \gamma_{j,m}\beta_{m} $$

and:

$$
\gamma_{j,m} \sim Lognormal(\mu_{m}, \sigma_{m})
$$

i.e. a draw from a Lognormal distribution.
$\gamma$ controls how much more or less risky a particular Location of a specific Setting Type is relative to the **average** for that Setting Type.
For example, $\gamma = 1$ is a setting which is as "risky" (or conducive to transmission) as the average location of that Setting Type.
$\gamma = 2$ would describe a setting that is twice as risky; $\gamma = 0.5$ would describe a setting that is half as risky etc.
Parameterisation and estimation of this parameter $\gamma$ is described further below.

### Estimating $\gamma$

In order to generate reasonable estimates of $\gamma$, we integrated a mathematical model of indoor air quality (as applied to infective virus concentrations) with a well-studied airborne infectious risk model (the Wells-Riley equation).
We then parameterised these models with estimates from the literature of 1) virus-specific parameters (with SARS-CoV-2 chosen as a case study) and 2) variation in ventilation rates across Locations belonging to the same Setting Type.
Together, these components enable us to produce estimates of the variation in riskiness for SARS-CoV-2 transmission between different Locations belonging to the same Setting Type.
Below we describe each of these components in turn.

#### Modelling the Concentration of Infective Virus in an Enclosed Space

Following the approach of [Blatchley et al, 2024](https://pubs.acs.org/doi/full/10.1021/acs.est.3c03026#), we model the concentration of pathogen in the air in an enclosed space via the following ODE:

$$
\nu \frac{dC(t)}{dt} =  - QC - k_DC \nu + I\pi 
$$

where $v$ is the space volume, $C$ is the infective virus concentration, $t$ is time, $Q$ is the flow rate of air into and out of the space, $k_D$ is the natural decay constant, $I$ is the number of infectious individuals in the space, and $\pi$ is the per capita (infective) virus emission rate.
The closed-form solution to this equation is:

$$
C =  \frac{I\pi}{\alpha \nu}(1 - e^{-\alpha t}) 
$$

where:

$$
\alpha = \frac{Q}{\nu} + k_D
$$

$\frac{Q}{\nu}$ is sometimes referred to as $A$ or the number of hair changes per hour.
$\alpha$ is therefore the number of equivalent air changes per hour taking into account both natural decay of the virus and ventilation driven removal of infective virus.

As $t \rightarrow\infty$, we get the steady state solution, which is the following:

$$
C_{ss} =  \frac{I\pi}{\alpha \nu}
$$

We note here also that for the purposes of the analyses presented here, we calculate the volume of the space as follows: $$
\nu = N D_mH
$$

where $N$ is the number of people occupying the space, $D_m$ is the density of people in the space (i.e. the number of people per m\^2) and $H$ is the height of the space.

#### Linking this Concentration of Infective Virus to an Infection Risk

We integrate this estimate of the concentration of infective virus in the room from the approach of [Blatchley et al, 2024](https://pubs.acs.org/doi/full/10.1021/acs.est.3c03026#) with the Wells-Riley equation, a simple and widely used model for the airbone transmission of infectious diseases.
The Wells-Riley equation takes the following form:

$$
P_{i}(t) =  1 - e^{-r C_{ss}Bt}
$$

where $P_i$ is the probability a person becomes infected conditional on them spending time $t$ in a room with $C_{ss}$ concentration of infective virus, $B$ is the volume rate at which air is breathed in, and $r$ is the risk parameter, the probability that a single inhaled infective virus will initiate an infection.

#### Estimating Variation in Riskiness Across Different Locations Belonging to the Same Setting Type

We carried out a sensitivity analysis to derive reasonable bounds for the $\gamma$ parameter described above.
We used the above equations in tandem with estimates of variation in the ventilation rates in different spaces and parameter estimates for SARS-CoV-2 from Blatchley et al (see Table \@ref(tab:1)) to explore how Location riskiness varied whilst varying:

1.  Infection Prevalence (i.e. the number of infectious individuals in the space).
2.  Air Changes Per Hour
3.  The Density of Individuals in the Space
4.  The Time Spent in the Space

Code and analyses to go here

Write something that because of our parameterisation using infection prevalence, that both the I term and the room vol term have setting sizes in them, which cancel out and so it's really density and the air changes per hour that drives things.

As we can see, the ratio is

| Parameter       | Value Used                     | Description                                                                                                                                                                         | Source                                                                                                                                                                                                                                                                                                                                                                                                                                  |
|-----------------|--------------------------------|-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| $\pi$           | 27 viruses per person per hour | Rates of (infective) virus emission from infected individuals.                                                                                                                      | [Blatchley et al, 2024](https://pubs.acs.org/doi/full/10.1021/acs.est.3c03026#) & [Buonanno et al, 2020](https://www.sciencedirect.com/science/article/pii/S0160412020320675#f0020)                                                                                                                                                                                                                                                     |
| $A_{school}$    | 0.1-2.9 air changes per hour   | Range of air-changes per hour considered for schools.                                                                                                                               | [Corsi et al, 2006](https://corsiaq.com/reports-etc/) - Section 3.1                                                                                                                                                                                                                                                                                                                                                                     |
| $D_{school}$    | 30 per 100m\^2                 | Assumed density of individuals in schools                                                                                                                                           | ANSI/ASHRAE Standard 62.1-2022 Table 6.1                                                                                                                                                                                                                                                                                                                                                                                                |
| $A_{workplace}$ | 0.22-4.84 air changes per hour | Range of air-changes per hour considered for workplaces.                                                                                                                            | [Corsi et al, 2006](https://corsiaq.com/reports-etc/) - Section 3.1                                                                                                                                                                                                                                                                                                                                                                     |
| $D_{workplace}$ | 10 per 100m\^2                 | Assumed density of individuals in workplaces                                                                                                                                        | ANSI/ASHRAE Standard 62.1-2022 Table 6.1                                                                                                                                                                                                                                                                                                                                                                                                |
| $A_{household}$ | 0.21-1.48 air changes per hour | Range of air-changes per hour considered for households.                                                                                                                            | [Corsi et al, 2006](https://corsiaq.com/reports-etc/) - Section 3.1                                                                                                                                                                                                                                                                                                                                                                     |
| $D_{household}$ | 5 per 100m\^2                  | Density of individuals in households                                                                                                                                                | Assumed                                                                                                                                                                                                                                                                                                                                                                                                                                 |
| $A_{leisure}$   | 0.22-4.84 air changes per hour | Range of air-changes per hour considered for leisure settings.                                                                                                                      | Assumed to be the same as workplaces.                                                                                                                                                                                                                                                                                                                                                                                                   |
| $D_{leisure}$   | 50 per 100m\^2                 | Density of individuals in leisure settings                                                                                                                                          | ANSI/ASHRAE Standard 62.1-2022 Table 6.1                                                                                                                                                                                                                                                                                                                                                                                                |
| $k_D$           | 0.64                           | Natural decay constant for SARS-CoV-2 in aerosols                                                                                                                                   | [Blatchley et al, 2024](https://pubs.acs.org/doi/full/10.1021/acs.est.3c03026#) & [van Doremalen et al, 2020](https://www.nejm.org/doi/full/10.1056/NEJMc2004973)                                                                                                                                                                                                                                                                       |
| $H$             | 2.5m                           | Height of the space                                                                                                                                                                 | Assumed                                                                                                                                                                                                                                                                                                                                                                                                                                 |
| $B$             | 0.45 m\^3 per hour             | Volume breathing rate i.e. the volume of air breathed in and out per hour. Assumes 15 breaths per minute and 500ml tidal volume to produce 0.45m\^3 volume breathing rate per hour. | [the internet - <https://www.verywellhealth.com/tidal-volume-5090250#>[:\~:text=What%20Is%20Tidal%20Volume%3F&text=Tidal%20volume%20(Vt%20or%20TV,mL%20for%20those%20assigned%20female](<https://www.verywellhealth.com/tidal-volume-5090250#>:\~:text=What%20Is%20Tidal%20Volume%3F&text=Tidal%20volume%20(Vt%20or%20TV,mL%20for%20those%20assigned%20female). and <https://my.clevelandclinic.org/health/articles/10881-vital-signs>] |
| $r$             | $1.37 × 10^{-2}$               | The probability that a single inhaled infective virus will initiate an infection.                                                                                                   | [Blatchley et al, 2024](https://pubs.acs.org/doi/full/10.1021/acs.est.3c03026#) & [Killingley et al, 2022](https://www.nature.com/articles/s41591-022-01780-9)                                                                                                                                                                                                                                                                          |

: Table 1: Parameter Estimates Used to Calculate Variation in Riskiness Across Locations Belonging to the Same Setting Type

# Assessing how various far UVC parameters shape another outcome

We considered two pathogen archetypes, SARS-CoV-2 and influenza.
The properties of SARS-CoV-2 were as follows.
The properties of influenze were as follows.
For each pathogen archetype, we considered the effect of far UVC efficacy and far UVC coverage on another outcome.
We varied far UVC efficacy in the following range.
We varied far UVC coverage in the following range with the following settings.
The results are presented in Figure \@ref(fig:placeholder-1) and Figure \@ref(fig:placeholder-2).

```{r placeholder-1, fig.cap="(ref:placeholder-1)"}
data.frame(x = 1:5, y = 1:5) |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

(ref:placeholder-1) Placeholder one.

```{r placeholder-2, fig.cap="(ref:placeholder-2)"}
data.frame(x = 1:5, y = 1:5) |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

(ref:placeholder-2) Placeholder two.

# Bibliography {.unnumbered}
