---
title: 'Update to "assessing far UVC interventions with an individual based infectious disease model"'
author:
  - Charlie Whittaker^[Imperial College London], Tom Brewer^[Imperial College London], Adam Howes^[Independent]
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{blueprint}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: REFERENCES.bib
editor_options: 
  markdown: 
    wrap: sentence
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 330,
  out.width = "90%"
)

```

```{r}
library(helios)
library(individual)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
theme_set(theme_minimal())
blueprint_colours <- colorRampPalette(c("#00AFFF", "#03113E"))(4)
exemplar_data <- readRDS("./exemplar_simulations.rds")
parameter_sweep_data <- readRDS("./parameter_sweep_data.rds")
```

# Overview

This document builds on [previous work](https://mrc-ide.github.io/helios/articles/blueprint.html) that introduced `helios`, an individual-based modelling framework for exploring the impact that installation of far UVC could have on the transmission and burden of respiratory infections.
In this document, we present a series of analyses evaluating the impact that installation of far UVC in non-household settings could have on the burden....

# Major Model Changes

Include relevant details about the model here, and changes from the first report:

## The Consistent Use of Data from the United States of America {#locations}

Add some subsections here about the data we now use. Can basically copy the structure of the first report.

## Location-Specific Variability in "Riskiness"

The concentration of pathogen in the air is modelled via the ODE:

$$
\nu \frac{dC(t)}{dt} =  - QC - k_DC \nu + I\pi 
$$

where $v$ is the room volume, $C$ is the infective virus concentration, $t$ is time, $Q$ is the flow rate of air into and out of the room, $k_D$ is the natural decay constant, $I$ is the number of infectious individuals in the room, and $\pi$ is the per capita (infective) virus emission rate.
The closed-form solution to this equation is:

$$
C =  \frac{I\pi}{\alpha \nu}(1 - e^{-\alpha t}) 
$$

where:

$$
\alpha = \frac{Q}{\nu} + k_D
$$

is the number of equivalent air changes per hour.
As $t \rightarrow\infty$, we get the steady state solution, which is the following:

$$
C_{ss} =  \frac{I\pi}{\alpha \nu}
$$

We integrate this estimate of the concentration of infective virus in the room with the Wells-Riley equation, a simple and widely used model for the airbone transmission of infectious diseases.
The Wells-Riley model takes the following form:

$$
P_{i}(t) =  1 - e^{-r C_{ss}Bt}
$$

where $P_i$ is the probability a person becomes infected conditional on them spending time $t$ in a room with $C_{ss}$ concentration of infective virus, $B$ is the volume rate at which air is breathed in, and $r$ is risk parameter, the probability that a single inhaled infective virus will initiate an infection.

# Assessing how various far UVC parameters shape another outcome

We considered two pathogen archetypes, SARS-CoV-2 and influenza.
The properties of SARS-CoV-2 were as follows.
The properties of influenze were as follows.
For each pathogen archetype, we considered the effect of far UVC efficacy and far UVC coverage on another outcome.
We varied far UVC efficacy in the following range.
We varied far UVC coverage in the following range with the following settings.
The results are presented in Figure \@ref(fig:placeholder-1) and Figure \@ref(fig:placeholder-2).

```{r placeholder-1, fig.cap="(ref:placeholder-1)"}
data.frame(x = 1:5, y = 1:5) |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

(ref:placeholder-1) Placeholder one.

```{r placeholder-2, fig.cap="(ref:placeholder-2)"}
data.frame(x = 1:5, y = 1:5) |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

(ref:placeholder-2) Placeholder two.

# Bibliography {.unnumbered}
