---
title: 'Update to "assessing far UVC interventions with an individual based infectious disease model"'
author:
  - Charlie Whittaker^[Imperial College London], Tom Brewer^[Imperial College London], Adam Howes^[Independent]
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{blueprint}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: REFERENCES.bib
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 330,
  out.width = "90%"
)

```

```{r}
library(helios)
library(individual)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
theme_set(theme_minimal())
blueprint_colours <- colorRampPalette(c("#00AFFF", "#03113E"))(4)
exemplar_data <- readRDS("./exemplar_simulations.rds")
parameter_sweep_data <- readRDS("./parameter_sweep_data.rds")
```

# Background

Include relevant details about the model here, and changes from the first report: [Assessing far UVC interventions with an individual based infectious disease model](https://mrc-ide.github.io/helios/articles/blueprint.html).

# Assessing how various far UVC parameters shape another outcome

We considered two pathogen archetypes, SARS-CoV-2 and influenza.
The properties of SARS-CoV-2 were as follows.
The properties of influenze were as follows.
For each pathogen archetype, we considered the effect of far UVC efficacy and far UVC coverage on another outcome.
We varied far UVC efficacy in the following range.
We varied far UVC coverage in the following range with the following settings.
The results are presented in Figure \@ref(fig:placeholder-1) and Figure \@ref(fig:placeholder-2).

```{r placeholder-1, fig.cap="(ref:placeholder-1)"}
data.frame(x = 1:5, y = 1:5) |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

(ref:placeholder-1) Placeholder one.

```{r placeholder-2, fig.cap="(ref:placeholder-2)"}
data.frame(x = 1:5, y = 1:5) |>
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  theme_minimal()
```

(ref:placeholder-2) Placeholder two.

# Bibliography {.unnumbered}
