---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction {#intro}

[`helios`](https://mrc-ide.github.io/helios/) is an R package which allows users to simulate the impact of [far UVC](https://www.nature.com/articles/s41598-018-21058-w) interventions in curtailing the spread of infectious disease outbreaks.

This vignette serves to demonstrate basic use of the package.
We start by loading the package, as well as other packages used in the vignette:

```{r setup}
library(helios)
library(individual)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
library(purrr)
theme_set(theme_minimal())
```

## Parameters {#param}

First, we define some constant parameters to be used in the model.
In this vignette we use the default options, aside from the `human_population` variable.
To facilitate reproducibility, we also set the seed for random number generation:

```{r}
parameters_list <- get_parameters(list(human_population = 5000, seed = 1))
```

The names of all parameters are as follows:

```{r}
names(parameters_list)
```

## Variables {#var}

Next, we define the model variables:

```{r}
variables_list <- create_variables(parameters_list)
names(variables_list)
```

Aside from `leisure`, each entry of `variables_list` is of the [`CategoricalVariable`](https://mrc-ide.github.io/individual/reference/CategoricalVariable.html) class:

```{r}
map(variables_list, class)
```

The `leisure` variable is a [`RagggedInteger`](https://mrc-ide.github.io/individual/reference/RaggedInteger.html) allowing its elements to have different lengths.

Some variables change over time (`disease_state`, `leisure`, `specific_leisure`), while other remain fixed (`age_class`, `workplace`, `school`, `household`).

### Disease states

The outbreak is implemented as individual-level SEIR [compartmental model](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology).

```{r}
(disease_states <- variables_list$disease_state$get_categories())
```

The number of individuals initially exposed is controlled by `parameters_list$number_initially_exposed`.

```{r}
parameters_list$number_initially_exposed
```

Initially, all other individuals are placed into the susceptible disease state:

```{r}
disease_state_counts <- purrr::map_vec(disease_states, function(x) variables_list$disease_state$get_size_of(values = x))

data.frame("State" = disease_states, "Count" = disease_state_counts) |>
  gt::gt()
```

### Age classes

There are three age classes: children, adults, and the elderly.
The age of each individual remains fixed throughout the simulation.

```{r}
parameters_list[c("initial_proportion_child", "initial_proportion_adult", "initial_proportion_elderly")]

age_classes <- variables_list$age_class$get_categories()
age_class_counts <- purrr::map_vec(age_classes, function(x) variables_list$age_class$get_size_of(values = x))

data.frame(age_classes, age_class_counts) |>
  mutate(age_classes = forcats::fct_relevel(age_classes, "child", "adult", "elderly")) |>
  ggplot(aes(x = age_classes, y = age_class_counts)) +
  geom_col() +
  labs(x = "Age class", y = "Count") +
  coord_flip()
```

### Schools

Each child attends a fixed school.
Additionally, a small number of adults attend schools as staff:

```{r}
schools <- variables_list$school$get_categories()
schools <- schools[schools != "0"]
school_sizes <- purrr::map_vec(schools, function(x) variables_list$school$get_size_of(values = x))

data.frame(school_sizes) |>
  ggplot(aes(x = school_sizes)) +
  geom_histogram() +
  labs(x = "School size", y = "Count")
```

### Workplaces

Each adult (aside from those who are school staff) attends a fixed workplace:

```{r}
workplaces <- variables_list$workplace$get_categories()
workplaces <- workplaces[workplaces != "0"]
workplace_sizes <- purrr::map_vec(workplaces, function(x) variables_list$workplace$get_size_of(values = x))

data.frame(workplace_sizes) |>
  ggplot(aes(x = log(workplace_sizes))) +
  geom_histogram() +
  labs(x = "log(Workplace size)", y = "Count")
```

### Households

Children, adults and elderly people are grouped into fixed households:

```{r}
households <- variables_list$household$get_categories()
household_sizes <- purrr::map_vec(households, function(x) variables_list$household$get_size_of(values = x))

table(household_sizes) |>
  data.frame() |>
  ggplot(aes(x = household_sizes, y = Freq)) +
  geom_col() +
  labs(x = "Household size", y = "Count")
```

The age distribution in households is as follows:

```{r}
household_df <- purrr::map_df(households, function(x) {
  indices <- variables_list$household$get_index_of(values = x)$to_vector()
  if(length(indices > 0)) data.frame("individual" = indices, "household" = as.numeric(x))
})

age_df <- purrr::map_df(age_classes, function(x) {
  indices <- variables_list$age_class$get_index_of(values = x)$to_vector()
  if(length(indices > 0)) data.frame("individual" = indices, "age_class" = x)
})

household_df |>
  left_join(age_df, by = "individual") |>
  group_by(household) |>
  summarise(
    child = sum(age_class == "child"),
    adult = sum(age_class == "adult"),
    elderly = sum(age_class == "elderly")
  ) |>
  group_by(child, adult, elderly) |>
  summarise(
    count = n()
  ) |>
  ungroup() |>
  arrange(desc(count)) |>
  head() |>
  gt::gt()
```


### Leisure

Each week, individuals attend leisure venues:

```{r}
leisure_places <- variables_list$leisure$get_values()
number_leisure_places <- sapply(leisure_places, function(x) sum(x > 0))

table(number_leisure_places) |>
  data.frame() |>
  ggplot(aes(x = number_leisure_places, y = Freq)) +
  geom_col() +
  labs(x = "Number of leisure places attended in a week", y = "Count")
```

## Events

Events govern the transition of individuals between disease states:

```{r}
events_list <- create_events(variables_list = variables_list, parameters_list = parameters_list)
names(events_list)
```

Each event is in the [`TargetedEvent`](https://mrc-ide.github.io/individual/reference/TargetedEvent.html) class:

```{r}
lapply(events_list, class)
```

## Render

We use a fixed number of time-steps to simulate over:

```{r}
timesteps <- round(parameters_list$simulation_time / parameters_list$dt)
```

The object `renderer` is of class [`Render`](https://mrc-ide.github.io/individual/reference/Render.html), and stores output from the model at each timestep:

```{r}
renderer <- individual::Render$new(timesteps)
class(renderer)
```

## Processes

Processes enable queing of events:

```{r}
processes_list <- create_processes(
  variables_list = variables_list,
  events_list = events_list,
  parameters_list = parameters_list,
  renderer = renderer
)

names(processes_list)
```

## Simulate

See [`individual::simulation_loop`](https://mrc-ide.github.io/individual/reference/simulation_loop.html):

```{r}
individual::simulation_loop(
  variables = variables_list,
  events = events_list,
  processes = processes_list,
  timesteps = timesteps
)
```

The output of the simulation is contained within `renderer`, and can be accessed using the `to_dataframe()` method:

```{r}
states <- renderer$to_dataframe()

states |>
  tidyr::pivot_longer(cols = ends_with("count"), names_to = "compartment", values_to = "value", names_pattern = "(.*)_count") |>
  mutate(
    compartment = forcats::fct_relevel(compartment, "S", "E", "I", "R")
  ) |>
  ggplot(aes(x = timestep, y = value, col = compartment)) +
  geom_line() +
  scale_color_manual(values = c("royalblue3", "firebrick3", "darkorchid3", "orange2")) +
  labs(x = "Time-step", y = "Count", col = "Compartment")
```
