---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{get-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction {#intro}

`helios` is an R package which allows users to simulate the impact of [far UVC](https://www.nature.com/articles/s41598-018-21058-w) interventions in curtailing the spread of infectious disease outbreaks.
In this vignette, we demonstrate basic use of the package.

```{r setup}
library(helios)
library(individual)
library(ggplot2)
library(dplyr)
library(purrr)
theme_set(theme_minimal())
```

## Parameters {#param}

First, define constant parameters.
In this vignette, we set the `human_population` to 1000.
Otherwise, we use the default options:

```{r}
parameters_list <- get_parameters(list(human_population = 1000, seed = 1))
```

The names of all parameters are as follows:

```{r}
names(parameters_list)
```

## Variables {#var}

Next, we define the model variables:

```{r}
variables_list <- create_variables(parameters_list)
names(variables_list)
```

Aside from `$leisure`, each entry of `variables_list` is of the [`CategoricalVariable`](https://mrc-ide.github.io/individual/reference/CategoricalVariable.html) class:

```{r}
map(variables_list, class)
```

Some variables change over time, while other remain fixed.

### Disease states

Outbreaks are implemented as individual-level SEIR [compartmental models](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology).

```{r}
(disease_states <- variables_list$disease_state$get_categories())
```

The number of individuals initially exposed is controlled by `parameters_list$number_initially_exposed`.

```{r}
disease_state_counts <- purrr::map_vec(disease_states, function(x) variables_list$disease_state$get_size_of(values = x))

data.frame("State" = disease_states, "Count" = disease_state_counts) |>
  gt::gt()
```

### Age classes

Age classes remain fixed and are as follows:

```{r}
(age_classes <- variables_list$age_class$get_categories())
age_class_counts <- purrr::map_vec(age_classes, function(x) variables_list$age_class$get_size_of(values = x))

data.frame(age_classes, age_class_counts) |>
  ggplot(aes(x = age_classes, y = age_class_counts)) +
  geom_col() +
  labs(x = "Age class", y = "Count") +
  coord_flip()
```

### Workplaces

Each adult attends a fixed workplace:

```{r}
workplaces <- variables_list$workplace$get_categories()
workplace_sizes <- purrr::map_vec(workplaces, function(x) variables_list$workplace$get_size_of(values = x))

data.frame(workplace_sizes) |>
  ggplot(aes(x = log(workplace_sizes))) +
  geom_histogram() +
  labs(x = "log(Workplace size)", y = "Count")
```

### Schools

Each child attends a fixed school:

```{r}
schools <- variables_list$school$get_categories()
school_sizes <- purrr::map_vec(schools, function(x) variables_list$school$get_size_of(values = x))

data.frame(school_sizes) |>
  ggplot(aes(x = log(school_sizes))) +
  geom_histogram() +
  labs(x = "log(School size)", y = "Count")
```

### Households

Children, adults and elderly people are grouped into fixed households:

```{r}
households <- variables_list$household$get_categories()
household_sizes <- purrr::map_vec(households, function(x) variables_list$household$get_size_of(values = x))

table(household_sizes) |>
  data.frame() |>
  ggplot(aes(x = household_sizes, y = Freq)) +
  geom_col() +
  labs(x = "Household size", y = "Count")
```

What is the age distribution in households?

```{r}
household_df <- purrr::map_df(households, function(x) {
  indices <- variables_list$household$get_index_of(values = x)$to_vector()
  if(length(indices > 0)) data.frame("individual" = indices, "household" = as.numeric(x))
})

age_df <- purrr::map_df(age_classes, function(x) {
  indices <- variables_list$age_class$get_index_of(values = x)$to_vector()
  if(length(indices > 0)) data.frame("individual" = indices, "age_class" = x)
})

household_df |>
  left_join(age_df, by = "individual") |>
  group_by(household) |>
  summarise(
    child = sum(age_class == "child"),
    adult = sum(age_class == "adult"),
    elderly = sum(age_class == "elderly")
  ) |>
  group_by(child, adult, elderly) |>
  summarise(
    count = n()
  ) |>
  ungroup() |>
  arrange(desc(count)) |>
  head() |>
  gt::gt()
```


### Leisure

Each week, individuals attend leisure venues:

```{r}
leisure_places <- variables_list$leisure$get_values()
number_leisure_places <- sapply(leisure_places, function(x) sum(x > 0))

table(number_leisure_places) |>
  data.frame() |>
  ggplot(aes(x = number_leisure_places, y = Freq)) +
  geom_col() +
  labs(x = "Number of leisure places attended in a week", y = "Count")
```

## Events

```{r}
events_list <- create_events(variables_list = variables_list, parameters_list = parameters_list)
names(events_list)
```

Each event is in the [`TargetedEvent`](https://mrc-ide.github.io/individual/reference/TargetedEvent.html) class:

```{r}
lapply(events_list, class)
```

## Render

We use a fixed number of time-steps to simulate over:

```{r}
timesteps <- round(parameters_list$simulation_time / parameters_list$dt)
```

The object `renderer` is of class [`Render`](https://mrc-ide.github.io/individual/reference/Render.html):

```{r}
renderer <- individual::Render$new(timesteps)
class(renderer)
```

## Processes

```{r}
processes_list <- create_processes(
  variables_list = variables_list,
  events_list = events_list,
  parameters_list = parameters_list,
  renderer = renderer
)

names(processes_list)
```

## Simulate

See [`individual::simulation_loop`](https://mrc-ide.github.io/individual/reference/simulation_loop.html):

```{r}
individual::simulation_loop(
  variables = variables_list,
  events = events_list,
  processes = processes_list,
  timesteps = timesteps
)
```

The output of the simulation is contained within `renderer`, and can be accessed using the `to_dataframe()` method:

```{r}
states <- renderer$to_dataframe()

states |>
  tidyr::pivot_longer(cols = ends_with("count"), names_to = "compartment", values_to = "value", names_pattern = "(.*)_count") |>
  mutate(
    compartment = forcats::fct_relevel(compartment, "S", "E", "I", "R")
  ) |>
  ggplot(aes(x = timestep, y = value, col = compartment)) +
  geom_line() +
  scale_color_manual(values = c("royalblue3", "firebrick3", "darkorchid3", "orange2")) +
  labs(x = "Time-step", y = "Count", col = "Compartment")
```
